---
title: "Homework 4"
author: "Shen Jialun 16307110030"
date: "10/28/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
cat("\014");rm(list=ls())
```

## 1-Ex5.15
We need to estimate
$$
\theta = \int_0^1\frac{e^{-x}}{1+x^2}dx
$$
Let
$$
g(x) = \frac{e^{-x}}{1+x^2}, \quad 0<x<1
$$
The importance funciton is
$$
f(x)  = \frac{e^{-x}}{1-e^{-1}},\quad 0<x<1
$$
the corresponding cdf is
$$
F(x) = \int_0^xf(x)dx = \frac{1-e^{-x}}{1-e^{-1}}
$$
the corresponding quantile function is
$$
F^{-1}(u) = -{\rm log}(1-u(1-e^{-1})),\quad 0\leq u\leq 1
$$

### (a)
If the integral range $(0,1)$ is divided into five equal-length subintervals,$I^{(a)}_j=((j-1)/5,j/5),j=1,2,...,5$, then on the $j^{th}$ subinterval, the density is
$$
f^{(a)}_j(x) = \frac{f(x)}{{\rm P}\Big(x\in I^{(a)}_j\Big)} = \frac{f(x)}{F(j/5)-F((j-1)/5)},\quad x\in I^{(a)}_j
$$

### (b)
If the integral range is divided into five pieces with quantiles of the importance function, i.e., the subintervals being
$$
I^{(b)}_j = \{x:a_{j-1}\leq x\leq a_j\},\quad j=1,2...,5
$$
where $a_j=F^{-1}(j/5), j=0,1,...,5$.

Then on the $j^{th}$ subinterval, variables are generated from the density
$$
f^{(b)}_j(x) = \frac{f(x)}{{\rm P}\Big(x\in I^{(b)}_j\Big)} = \frac{5e^{-x}}{1-e^{-1}},\quad x\in I^{(b)}_j
$$

```{r}
M <- 10000
theta.hat <- se <- numeric(3)
g <- function(x){
  exp(-x - log(1+x^2)) * (x>0) * (x<1)
}
f <- function(x){
  exp(-x) / (1 - exp(-1))
}
F.cdf <- function(x){
  (1 - exp(-x)) / (1 - exp(-1))
}
F.quantile <- function(u){
  - log(1- u * (1 - exp(-1)))
}
#using importance sampling
u <- runif(M)
x <- F.quantile(u) #inverse transform method
fg <- g(x) / f(x)
theta.hat[1] <- mean(fg)
se[1] <- sd(fg)
#using stratified importance sampling with subinterval Ia
k <- 5
m <- M/k
T1 <- var1 <- numeric(5)
a <- seq(0, 1, length.out = 6)
for (j in 1:5){
  ai <- a[j] #a_{i-1}
  aii <- a[j+1] #a_i
  fj <- function(x){f(x) / (F.cdf(j/5) - F.cdf((j-1)/5))}
  cnt <- 0
  fg <- numeric(m)
  while (cnt < m){
    u <- runif(1)
    x <- F.quantile(u)
    if ((ai < x) & (x < aii)){
      cnt <- cnt+1
      fg[cnt] <- g(x)/fj(x)
    }
  }
  T1[j] <- mean(fg)
  var1[j] <- var(fg)
}
theta.hat[2] <- sum(T1)
se[2] <- sqrt(1/m * sum(var1))
#using stratified importance sampling with subinterval Ib
T2 <- var2 <- numeric(5)
dim(a) <- length(a)
b <- apply(a, 1, F.quantile)
fj <- function(x){5 * f(x)}
for (j in 1:5){
  ai <- b[j] #a_{i-1}
  aii <- b[j+1] #a_i
  cnt <- 0
  fg <- numeric(m)
  while (cnt < m){
    u <- runif(1)
    x <- F.quantile(u)
    if ((ai < x) & (x < aii)){
      cnt <- cnt+1
      fg[cnt] <- g(x)/fj(x)
    }
  }
  T2[j] <- mean(fg)
  var2[j] <- var(fg)
}
theta.hat[3] <- sum(T2)
se[3] <- sqrt(1/m * sum(var2))
theta.hat
se
```

## 2-Ex6.1
The pdf of a standard Cauchy distribution is
$$
f(x) = \frac{1}{\pi (1+x^2)}
$$

```{r}
n <- 20
K <- n/2 - 1
m <- 1000
mse <- matrix(0, K, 2)
trimmed.mse <- function(n, m, k){
  #MC est for k-level trimmed mean of standard Cauchy distribution
  tmean <- numeric(m)
  for (i in 1:m){
    x <- sort(rcauchy(n))
    tmean[i] <- sum(x[(k+1):(n-k)]) / (n-2*k)
  }
  #theta <- mean(tmean)#x
  theta <- median(tmean)
  mse.hat <- mean((tmean - theta)^2)
  se.mse <- sqrt(mean((tmean - theta)^2))/sqrt(m)
  return (c(mse.hat, se.mse))
}
for (k in 1:K){
  mse[k, 1:2] <- trimmed.mse(n=n, m=m, k=k)
}
mse.table <- data.frame(seq(1,9), mse)
names(mse.table) <- c('k', 'mse of k-level trimmed mean', 'standard error')
mse.table
```

## 3
__Proof:__

Suppose $X_1,X_2,...,X_n$ are $n$ random samples independently generated from standard normal distribution. Sort them in increasing order, and we obtain $X_{(1)}\leq X_{(2)}\leq ...\leq X_{(n)}$. The $k$-level trimmed mean is
$$
\bar X_{[-k]} = \frac{1}{n-2k}\sum_{i=k+1}^{n-k}X_{(i)}
$$
The expectation of $\bar X_{[-k]}$ is
$$
\begin{aligned}
E(\bar X_{[-k]}) &= E\bigg(\frac{1}{n-2k}\sum_{i=k+1}^{n-k}X_{(i)}\bigg)\\
&= \frac{1}{n-2k}E\bigg(\sum_{i=k+1}^{n-k}X_{(i)}\bigg)\\
&= \frac{1}{n-2k}(n-2k)E(X)\\
&= \mu \\
&= 0
\end{aligned}
$$

$\Box$
